Think deep for reef resilience

### 2025-08-16 paper/paper_draft.qmd
Reinsert contextual detail around clustering
Correct typo

The time series clustering was applied over the time frame between 2030 and 2060. Spatially,
the analyses were conducted for each GBR management area and at the scale of each reefal
bioregion (defined in GBRMPA [-@GBRMPA2024]). Analyses were also performed at the
whole-of-GBR scale but were found to lead to identical results **\[TI: Ben, compared to
what, the bioregion or the management area results?\]** and so are not shown here. To
compare time series clustering results across GCMs, management area scale clustering results
are used to preserve coarse scale latitudinal differences between regions, while maintaining
ease of comparison and visualisation. These comparisons include quantification of
differences between low and high cover clusters across management regions under each GCM
using the CID between cluster median time series. To assess how different reef
characteristics influence time series clustering at a finer spatial scale, bioregion results
are used to capture patterns occurring within individual clusters of reefs. Results of all
analyses at all clustering scales can be found at Grier and Iwanaga [-@grier2025] for those
interested.

@@

These findings support the hypothesis that reefs at greater depths will exhibit
characteristics of refugia and more prolonged structural resilience under climate change
compared to shallower reefs. Additionally, Lui, Callaghan and Baldock [-@liu2024] suggest
that rubble instability is inversely correlated with reef depth, possibly limiting the
juvenile recruitment capacity of shallow reefs. A caveat to these depth-resilience findings
is the quality of the bathymetry data used to calculate median reef depth. As detailed
previously, estimates of median reef depth are derived from satellite imagery (at $10m^2$
resolution) which can be subject to high degrees of uncertainty, both in terms of the
estimated depth at for each pixel as well as their alignment to the established GBRMPA reef
outlines.

>>>>

The time series clustering was applied over the time frame between 2030 and 2060. The
selected time frame prevents the classification from being overly influenced by any steep
declines in the initial years of simulations, and avoids the post-2060 time period where
reefs are projected to be under severe stress and in decline [@bozec2025]. Spatially, the
analyses were conducted for each GBR management area and at the scale of each reefal
bioregion (defined in GBRMPA [-@GBRMPA2024]). Analyses were also performed at the
whole-of-GBR scale but were found to lead to identical results **\[TI: Ben, compared to
what, the bioregion or the management area results?\]** and so are not shown here. To
compare time series clustering results across GCMs, management area scale clustering results
are used to preserve coarse scale latitudinal differences between regions, while maintaining
ease of comparison and visualisation. These comparisons include quantification of
differences between low and high cover clusters across management regions under each GCM
using the CID between cluster median time series. To assess how different reef
characteristics influence time series clustering at a finer spatial scale, bioregion results
are used to capture patterns occurring within individual clusters of reefs. Results of all
analyses at all clustering scales can be found at Grier and Iwanaga [-@grier2025] for those
interested.

@@

These findings support the hypothesis that reefs at greater depths will exhibit
characteristics of refugia and more prolonged structural resilience under climate change
compared to shallower reefs. Additionally, Liu, Callaghan and Baldock [-@liu2024] suggest
that rubble instability is inversely correlated with reef depth, possibly limiting the
juvenile recruitment capacity of shallow reefs. A caveat to these depth-resilience findings
is the quality of the bathymetry data used to calculate median reef depth. As detailed
previously, estimates of median reef depth are derived from satellite imagery (at $10m^2$
resolution) which can be subject to high degrees of uncertainty, both in terms of the
estimated depth at for each pixel as well as their alignment to the established GBRMPA reef
outlines.
